---
title: "BTAP-MS"
author: "William E Fondrie"
date: "5/2/2018"
output: 
    github_document:
        pandoc_args: --webtex
---

# Introduction  

# Methods  

## Setup  
```{r knitrSetup, echo=F, include=FALSE}
knitr::opts_chunk$set(warning = F, message = F)
```  

```{r loadPkgs, results='hide'}
library(tidyverse)
library(broom)
library(ggridges)
library(knitr) # for the kable() function
library(seqinr) # calculate molecular weights

set.seed(190875) # to ensure reproducibility

# load auxillary functions
source("R/ggplotTheme.R") # Plot themes
full <- 7
half <- 3.33

# Set ggplot2 theme
theme_set(coolTheme)
```


## Load Skyline Results  

Precursor quantitation data was imported from the Skyline (exported "Transition Results" report) `skyline_results.csv`. Proteins were summarized as the sum of all precursor intensities for the protein. The `Replicate` column indicates the bait and concentration in the following form `...[bait]-Rep1-[3^x nM]`. Thus, `...LRP1B-Rep1-1` indicates that the GST-LRP1B-ICD was used as bait at a concentration of 3 nM. Samples with the GST suffix indicate that only GST was used.
```{r loadData}
prot <- read_csv("data/skyline_results.csv") %>%
    group_by(Protein, Replicate) %>%
    summarize(intensity = sum(Area, na.rm = T)) %>%
    ungroup() %>%
    mutate(bait = paste0("GST", str_match(Replicate, "BTAP(-.+-)Rep")[ , 2], "ICD"),
           bait = str_replace(bait, "LRP-", "LRP1-"),
           conc = str_match(Replicate, "Rep1-(.*?)(-|_|$)")[ , 2],
           conc = ifelse(conc == "GST", 0, 3^as.numeric(conc)))

# Retrieve protein info
fasta <- read.fasta("data/SwissProt-Human_v2018-02_GST.fasta", seqtype = "AA", 
                    as.string =T, set.attributes = T)

fastaKey <- tibble(Protein = names(fasta), sequence = unlist(fasta), 
                   annotation = map_chr(fasta, ~ attributes(.)$Annot))

protInfo <- prot %>%
    group_by(Protein) %>%
    summarize(accession = str_match(Protein, "\\|(.*)\\|") [1 , 2]) %>%
    left_join(fastaKey) %>%
    group_by(Protein) %>%
    mutate(MW = pmw(unlist(str_split(sequence, ""))) / 1000, #kDa
           gn = str_match(annotation, "GN=(.*?) ")[ , 2]) %>%
    select(Protein, accession, gn, MW) # only the cols we'll use

# Median normalization
globalMedian <- log2(median(prot$intensity))

lfq <- prot %>% 
    group_by(Replicate) %>% 
    mutate(medianIntensity = log2(median(intensity)),
           lfq = 2^(log2(intensity) - medianIntensity + globalMedian))

```

The number of proteins passing this stage of analysis is:  
```{r numProt1}
prot %>% 
  group_by(bait) %>%
  summarize(proteins = length(unique(Protein))) %>%
  kable()
```

## Transforming Intensities  

The measured intensities indicate the amount of unbound protein in each sample. To convert this to be proportional to the amount of bound protein, which is needed obtain familiar binding curve shapes, the maximum LFQ intensity for each protein was subtracted and the signed was inversed.
```{r lfqTransform}
lfqModInput <- lfq %>%
  group_by(Protein, bait) %>%
  mutate(response =  max(lfq) - lfq)
```

## Nonlinear Modeling  

Each protein was fit to the 1:1 equilibrium binding isotherm using nonlinear least-squares regression. The binding isotherm model takes the form of:  

$$ R = \frac{ [B]_t * R_{max} }{ [B]_t + K_d } $$  

Where given $R$ (the response) and $[B]_t$ (the total bait concentration), we fit the curves to estimate $R_{max}$ (the estimated maximal response) and $K_d$ (the equilibrium dissociation constant). 

```{r modeling}
mods <- lfqModInput %>%
  group_by(Protein, bait) %>%
  filter(!is.na(response)) %>%
  do(models = nls(response ~ (Rmax * conc) / (Kd + conc),
                  data = .,
                  start = list(Kd = 100, Rmax = 4e+07),
                  control = list(maxiter = 200,
                                 warnOnly = T)))

# Retrieve information about model fits
fitInfo <- mods %>% glance(models)

kable(fitInfo[1:5, ])

# Retrieve model parameter values
fitVals <- mods %>% 
  tidy(models) %>%
  mutate(CV = std.error / estimate * 100)

kable(fitVals[1:5, ])
```

Because many of the proteins we measured do not actually interact with the bait proteins, there are a large number of model fits that failed to converge. These were removed from consideration.

```{r removeNotConverged}
fits <- fitInfo %>%
  filter(isConv) %>%
  select(Protein, bait) %>%
  left_join(fitVals)
```

And lastly, we reshape the parameter values to a wide format.  

```{r paramTable}
RmaxTbl <- fits %>%
  filter(term == "Rmax") %>%
  select(Protein, bait, estimate, CV) %>%
  rename(Rmax = estimate, Rmax_CV = CV)

fitTbl <- fits %>%
  filter(term == "Kd") %>%
  left_join(RmaxTbl)

kable(fitTbl[1:5, ])
```

The number of proteins passing this stage of analysis is:  
```{r numProt3}
fitTbl %>% 
  group_by(bait) %>%
  summarize(proteins = length(unique(Protein))) %>%
  kable()
```

## Protein Concentration Estimation  

For the binding isotherm equation above to be valid, we must make the assumption that the conentration of prey protein, $[P]_t$ is much less than the $K_d$. In an effort to verify this assumption, the prey protein concentrations were crudely estimated using the "Total Protein Approach." This approach uses the following estimation to calculate the relative protein concentration in a shotgun proteomics study:    

$$ \frac{Protein~Mass}{Total~Protein~Mass} \approx \frac{Protein~MS~Signal}{Total~MS~Signal}$$  

Thus, given the mass spec signal of a protein, the total mass spec signal (the sum of intensities for a run) we can estimate the relative contributation of a single protein to the total protein mass analyzed. Because the experiment was performed at a total protein conentration of 1 ug/uL, we can then calculate the individual protein concentrations using their molecular weights. 


```{r TPAcalc}
tpaQuan <- lfqModInput %>%
  group_by(bait, conc) %>%
  left_join(protInfo) %>%
  mutate(tpa = lfq / sum(lfq, na.rm = T) / (MW * 1000),
         preyConc = tpa * 10^9,
         logPreyConc = log10(preyConc))

titer <- tpaQuan %>%
  group_by(Protein) %>%
  summarize(maxConc = max(preyConc, na.rm = T))
```

## Filtering For Sufficient Model Fits

While some estimated $K_d$ values are physically impossible, such as those below zero, others are outside of the range that this experiment was designed to measure. Because the bait concentrations used were between 1 and 2187 nM, we filtered for $K_d$ between 1 nM and 1000 nM. Additionally, higher coefficients of variation (CV) are indicative of poor model fits so we use it as an additional filter.  
```{r filterFits}
interactors <- fitTbl %>%
  left_join(titer) %>%
  filter(estimate > 1, 
         estimate < 1000,
         estimate > maxConc * 10) %>%
  arrange(CV) 
```

The number of proteins passing this stage of analysis is:  
```{r numProt4}
interactors %>% 
  group_by(bait) %>%
  summarize(proteins = length(unique(Protein))) %>%
  kable()
```

## Final interactor list
```{r final filter}
highConfCut <- quantile(interactors$CV, 0.05)

highConf <- interactors %>%
    ungroup() %>%
    filter(CV <= highConfCut) %>%
    arrange(CV) 

highConf %>% 
  group_by(bait) %>%
  summarize(proteins = length(unique(Protein))) %>%
  kable()

a <- filter(highConf, bait == "GST-LRP-ICD") %>%
    mutate(gn = str_match(Protein, "\\|.*\\|(.*)?_")[ , 2])
```


# Results 
## TPA Estimation Plots
```{r TPAplots}
boxPlot <- tpaQuan %>%
  ggplot(aes(x = as.factor(conc), y = logPreyConc)) +
  geom_violin(fill = ggColors(3)[3]) +
  geom_boxplot(fill = "grey", width = 0.2, outlier.shape = NA) +
  facet_wrap(~ bait, ncol = 1) +
  ylab(expression("log"[10]*"[Protein (nM)]")) +
  xlab("Bait (nM)")

savePlot(boxPlot, w = full/2, h = 4)

overallDist <- titer %>%
  ggplot(aes(x = log10(maxConc))) +
  geom_density(fill = ggColors(3)[3]) +
  xlab(expression("log"[10]*"[Protein (nM)]")) +
  ylab("Density")
savePlot(overallDist, w = full/2, h = 2)


# some concentration stats
titer %>% 
    summarize(Mean = mean(log10(maxConc)),
              `Mean (nM)` = 10^Mean,
              `Std Dev` = sd(log10(maxConc)),
              total = length(Protein),
              `< 1` = sum(maxConc < 1) / total * 100,
              `< 10` = sum(maxConc < 10) / total * 100,
              `< 100` = sum(maxConc < 100)/ total * 100) %>%
    kable

# park stats
tpaQuan %>%
    left_join(protInfo) %>%
    filter(gn == "PARK7") %>%
    ungroup() %>%
    summarize(`mean (nM)` = mean(preyConc),
              `se` = sd(preyConc) / sqrt(length(preyConc))) %>%
    kable


barPlot <- titer %>%
  summarize(total = length(Protein),
            `< 1` = sum(maxConc < 1) / total * 100,
            `< 10` = sum(maxConc < 10) / total * 100,
            `< 100` = sum(maxConc < 100)/ total * 100) %>%
  gather(conc, percent, starts_with("<")) %>%
  ggplot(aes(x = conc, y = percent, fill = conc)) +
  geom_col(position = "dodge", color = "black") +
  #geom_errorbar(aes(ymax = avg + ci, ymin = avg - ci), width = 0.25) +
  xlab("Protein (nM)") +
  ylab("Number of Proteins (%)") +
  theme(legend.position = "none")

savePlot(barPlot, w = full/2, h = 2)
```

## Evaluating Model Fit Plots
```{r fitEval}
fitScatter <- interactors %>% 
  ggplot(aes(x = log10(estimate), y = CV, color = bait)) +
  geom_point() +
  geom_hline(yintercept = highConfCut, linetype = "dashed") + 
  scale_color_discrete(name = "Bait") +
  theme(legend.position = c(0, 1),
        legend.justification = c(0, 1),
        legend.background = element_rect(color = "black"),
        legend.title.align = 0.5,
        legend.key.height = unit(0.5, "lines")) +
  xlab(expression("Log"[10]*"[K"[d]~" (nM)]")) +
  ylab("Coefficient of Variation (%)")

savePlot(fitScatter, w = half, h = half)

cvDens <- interactors %>%
  ggplot(aes(x = CV, fill = bait)) +
  geom_histogram() +
  geom_vline(xintercept = highConfCut, linetype = "dashed") +
  facet_wrap(~bait, ncol = 1) +
  theme(legend.position = "none") +
  ylab("Number of Proteins") +
  xlab("Coefficient of Variation (%)")
  
savePlot(cvDens, w = half, h = half)
```

## Binding Curve Plots
```{r plotExamples}
plotIsotherm <- function(dat, responseDat = lfqModInput, logScale = F, ncol = 3) {
  
    # Function to create binding curves
    curveFun <- function(Protein, Rmax, estimate, CV, gn, bait,..., conc) {
        response <- (Rmax * conc) / (estimate + conc)
        tibble(response = response,
               conc = conc,
               Protein = Protein,
               CV = CV,
               gn = gn,
               estimate = estimate,
               bait = bait)
    }
    
    
    if(logScale) {
      x <- 10^(seq(0,log10(2187), length.out = 100))
    } else {
      x <- seq(0, 2187, length.out = 100)
    }
    
    # create curve data
    curveDat <- dat %>%
        select(Protein, estimate, Rmax, CV, gn, bait)%>%
        pmap_df(curveFun, conc = x) %>%
        ungroup() %>%
        mutate(gn = fct_reorder2(gn, as.numeric(bait), estimate))
    
    # Add response data
    datR <- dat %>% 
        select(Protein, bait, estimate, Rmax, CV, gn) %>%
        left_join(responseDat) %>%
        ungroup() %>%
        mutate(gn = fct_reorder2(gn, as.numeric(bait), estimate))

    # make plot
    p <- datR %>%
        ggplot(aes(x = conc, y = response * 1e-7)) +
        geom_line(aes(color = bait), data = curveDat) +#, color = ggColors(3)[3]) +
        geom_point(size = 0.5) +
        facet_wrap(~gn, ncol = ncol, scales = "free_y") +
        ylab(expression("Response (10"^7*")")) +
        scale_color_discrete(name = "Bait")
      
    
    if(logScale){ 
        p <- p + scale_x_log10(breaks = c(1, 10, 100, 1000)) 
    } else {
        p <- p + scale_x_continuous(breaks = c(0, 1000, 2000))
    } 
    
    if(length(unique(dat$bait)) == 2) {
        p <- p + xlab("Bait (nM)") 
        
    } else {
        p <- p + xlab(paste0(dat$bait[1], " (nM)"))
    }
    
    p
}


allCurves <- highConf %>%
    left_join(protInfo) %>%
    plotIsotherm(logScale = T, ncol = 3) +
    theme(legend.position = c(0.75, 0),
          legend.justification = c(0, 0),
          legend.key.size = unit(0.5, "lines"),
          legend.background = element_rect(color = "black")) 
savePlot(allCurves, w = 2 * full / 3, h = 5)

```

## Plot Calculated Affinities and Error Bars
```{r morePlots}
exBar <- highConf %>%
    left_join(protInfo) %>%
    ungroup() %>%
    mutate(gn = fct_reorder2(gn, 
                             fct_rev(as.factor(bait)), 
                             estimate, 
                             .desc = F),
           affinity = ifelse(estimate >= 50, "low", "high")) %>% 
    ggplot(aes(x = gn, y = estimate, color = bait)) +
    geom_pointrange(aes(ymin = estimate - std.error, ymax = estimate + std.error),
                  width = 1, fatten = 2) +
    theme(legend.position = "none",
          axis.title.y = element_blank()) +
    facet_wrap(~bait, scales = "free_y", drop = T, ncol = 1) +
    ylab(expression("K"[d]~"(nM)")) +
    coord_flip()
savePlot(exBar, w = full / 3, h = 5)
```

## Info on specific proteins
```{r}
# lrp1 biogrid
bioGridLrp1 <- read_tsv("data/bioGrid_lrp1.txt")
lrp1Genes <- unique(c(bioGridLrp1$`Official Symbol Interactor A`, 
                      bioGridLrp1$`Official Symbol Interactor B`))

interactors %>%
    left_join(protInfo) %>%
    filter(gn %in% lrp1Genes,
           bait == "GST-LRP1-ICD") %>%
    kable

# lrp1b biogrid
bioGridLrp1b <- read_tsv("data/bioGrid_lrp1b.txt")
lrp1bGenes <- unique(c(bioGridLrp1b$`Official Symbol Interactor A`, 
                      bioGridLrp1b$`Official Symbol Interactor B`))

interactors %>%
    left_join(protInfo) %>%
    filter(gn %in% lrp1bGenes,
           bait == "GST-LRP1B-ICD") %>%
    kable


# all high conf
highConf %>%
    left_join(protInfo) %>%
    kable
```

## Write tables
```{r}
# to be added
```



# Session Info
```{r, results = "as-is"}
devtools::session_info()
```



